---
title: "Google Capstone Project"
output: github_document
date: "2024-01-15"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
chooseCRANmirror(graphics=FALSE, ind=1)

```

# Is it a good time to start a car sharing company in 2022?

## Background

A private investment would like to explore the opportunity of starting a
new car sharing/ rental company in the UK and compete with other
existing market holder such as ZipCar, Co-Wheel.

## Objective:

-   What car body type should be in the fleet?
-   What engine type of car should be considered?
-   Is 2022 a good time?
-   Potentially which city/ cities to launch if possible?

## Metric to use: Analyse data and find out:

-   Popular UK car
-   Cost to run and maintenance for different type of car
-   City with highest hotel booking rate

## Data used:

-   First time registered vehicle in the UK Office of National
    Statistics, Primary Data . For data [click
    here](https://www.gov.uk/government/statistical-data-sets/vehicle-licensing-statistics-data-tables)
-   Consumer Price Index for new car Office of National Statistics,
    Primary Data. For data [click
    here](https://www.ons.gov.uk/economy/inflationandpriceindices/timeseries/cjxn/mm23)
-   Non-domestic energy price Gov.UK, Primary Data. For data [click
    here](https://www.gov.uk/government/statistical-data-sets/gas-and-electricity-prices-in-the-non-domestic-sector)
-   UK petrol price Gov.UK, Primary Data. For data [click
    here](https://www.gov.uk/government/statistical-data-sets/gas-and-electricity-prices-in-the-non-domestic-sector)

All data are downloaded and store locally in author desktop

## Analysis

RStudio is used to clean and analysis the result Tabule used to create
dashboard and data visualisation

-   Total car sales in last 10 years
-   Top 10 best-selling car general model in 2021-2022Q1
-   Top 10 best-selling car specific model in 2021-2022Q1
-   CPI index of new car
-   UK Petrol and Electricity price

## Code

Install and load library

```{r import}
install.packages("tidyverse")
install.packages("zoo")
install.packages("ggplot2")
library(tidyverse)
library(zoo)
library(ggplot2)
library(readxl)
```

tidyverse is used for data cleaning ggplot2 is used for plotting grpahs
Load data into RStudio

```{r load data}
first_time_vehicle = read.csv("Dataset/First_time_register_vehicle.csv")
weekly_road_fuel_price = read.csv("Dataset/weekly_road_fuel_price.csv")
electric_price = read_xlsx("Dataset/table_341.xlsx", sheet =3, range = "A7:P80")
CPI_Data = read.csv("Dataset/New car CPI index.csv")
```

### Data cleaning for first time vehicle

```{r }
##rename the column
first_time_vehicle = first_time_vehicle %>%
  rename_with(~ gsub('X', '', .x))

##Filter and select relavent data
filter_data = first_time_vehicle %>%
  select("BodyType":"2010Q4", -"2022Q1") %>%
  filter(BodyType == "Cars")

##Tranpose data as date in different columns which will be hard to group data by date
tranpose_data = filter_data %>%
  pivot_longer(
    cols = starts_with("20"), 
    names_to = "YrQr", 
    values_to = "Sales",
    values_drop_na = TRUE)
```

Find out overall car sales since 2011

```{r eval=FALSE}
Yr_only_sale = tranpose_data%>%
  mutate(Year_only = format(as.Date(as.yearqtr(tranpose_data$YrQr)), "%Y")) %>%
  group_by(Year_only) %>%
  summarise(Annual_total_car_sale= sum(Sales))
```

Find out the best selling car since 2011

```{r Best selling car since 2011}
car_Model_sale_from_2011 = tranpose_data %>%
  group_by(Make) %>%
  group_by(GenModel) %>%
  summarise(Total_model_sales_2021to2011 = sum (Sales)) %>%
  arrange(-Total_model_sales_2021to2011)

head(car_Model_sale_from_2011)
```

Find out the best selling car from 2021 till 2022 in terms of General
Model.

```{r Best selling car 2021 to 2022}
filter_data_2021 = first_time_vehicle %>%
  select("BodyType":"2021Q1") %>%
  filter(BodyType == "Cars", !grepl("*MISSING", GenModel)) %>%
  mutate (Tota_sales = `2021Q1` + `2021Q2` + `2021Q3` + `2021Q4` + `2022Q1`, .keep = c("unused")) %>%
  group_by(GenModel) %>%
  summarise(Total_sales_by_genmodel = sum(Tota_sales)) %>%
  arrange(-Total_sales_by_genmodel)

head(filter_data_2021)
```

Find out the best selling car from 2021 till 2022 in terms of Model.

```{r best selling car from 2021 till 2022 in terms of Model}
Model_2021_22_sales = first_time_vehicle %>%
  select("BodyType":"2021Q1") %>%
  filter(BodyType == "Cars", !grepl("*MISSING", GenModel)) %>%
  mutate (Tota_sales = `2021Q1` + `2021Q2` + `2021Q3` + `2021Q4` + `2022Q1`, .keep = c("unused")) %>%
  group_by(Model) %>%
  #arrange(-Tota_sales)
  summarise(Total_sales_by_model = sum(Tota_sales)) %>%
  arrange(-Total_sales_by_model)

head(Model_2021_22_sales)
```

Annual electric price

```{r}
electric_price = electric_price %>%
  select(Year,Quarter,`Electricity: Average (Pence per kWh)`)%>%
  group_by(Year)%>%
  summarise(Annual_electricity = mean(`Electricity: Average (Pence per kWh)`))
```

CPI new car

```{r}
CPI_Data = read.csv("Dataset/New car CPI index.csv")
CPI_Data = CPI_Data[8:33,]
CPI_Data_year = CPI_Data %>%
  rename(Year = Title,  CPI_Index = CPI.INDEX.07.1.1A...NEW.CARS.2015.100)
CPI_Data_year$CPI_Index =as.numeric( CPI_Data_year$CPI_Index)
CPI_Data_year$Year =as.numeric(CPI_Data_year$Year)

head(CPI_Data_year)
```

Petrol pump price

```{r}
colnames(weekly_road_fuel_price) = c (weekly_road_fuel_price[1:1,])
weekly_road_fuel_price$`Pump price in pence/litre` = as.numeric(weekly_road_fuel_price$`Pump price in pence/litre`)

filtered_fuel = weekly_road_fuel_price[1:2]
filtered_fuel = filtered_fuel[-c(1,2),]
rownames(filtered_fuel) <- NULL

filtered_fuel = filtered_fuel %>%
  mutate(Monthly_price = format(as.Date(filtered_fuel$`Weekly Prices time series`, "%d/%m/%Y"), "%Y")) %>%
  filter(Monthly_price >= 2012) %>%
  relocate(Monthly_price, .before = `Pump price in pence/litre`) %>%
  group_by(Monthly_price) %>%
  summarise(Pump_price = mean(`Pump price in pence/litre`))
head(filtered_fuel)
```

## Plotting graph using ggplot2

Car sales

```{r eval=FALSE}
Yr_only_sale_1k=Yr_only_sale
Yr_only_sale_1k[,2]= Yr_only_sale_1k[,2]/1000
transform(Yr_only_sale_1k,Year_only= as.numeric(Year_only))
Yr_only_sale_1k$Year_only = as.numeric(Yr_only_sale_1k$Year_only)
```

```{r eval=FALSE}
annual_car_sale = ggplot(Yr_only_sale_1k) + 
  geom_line(mapping = aes(x=Year_only, y =(Annual_total_car_sale)), linetype = 1, color = "light blue", size =2, lineend = "round") +
  ylab("Annual Car Sales (thousands number)") + xlab("Year") +
  theme_bw() + theme(legend.title = element_blank())+
  geom_text(aes(label=ifelse(Year_only==2021, Annual_total_car_sale," "), x=Year_only, y =Annual_total_car_sale), hjust=0.5, vjust=1.5, fontface ="bold")
annual_car_sale
```

```{r}
##by genmodel 
top_10_car = filter_data_2021 [1:10, ]
top_10_genmodel_2021 = ggplot(top_10_car, aes(x=GenModel, y=Total_sales_by_genmodel, fill =GenModel) )+
  geom_col( alpha =2, size =0.5, color = "white")+
  theme_minimal()+
  xlab("General Model Name") + ylab("Sales (number)")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.6))+
  ylim(0,52000) +
  geom_text(aes(label=ifelse(GenModel=="VAUXHALL CORSA"|GenModel=="TESLA MODEL 3"| GenModel =="TOYOTA YARIS", Total_sales_by_genmodel,""), x=GenModel, y =Total_sales_by_genmodel), hjust=0.5, vjust=-0.2, fontface ="bold")
top_10_genmodel_2021

##by model
top_10_model = Model_2021_22_sales [1:10, ]

First_model = top_10_model[1,1]
Second_model = top_10_model[2,1]
Third_model = top_10_model[3,1]

top_10_model_2021 = ggplot(top_10_model, aes(x=Model, y=Total_sales_by_model, fill =Model) )+
  geom_col( alpha =2, size =0.5, color = "white")+
  theme_minimal()+
  xlab("Model Name") + ylab("Sales (number)")+
  theme(axis.text.x  = element_text(angle=45, vjust=1, hjust = 1))+
  ylim(0,30000) +
  geom_text(aes(label=ifelse(Model %in% Second_model| Model %in% First_model|Model %in% Third_model, Total_sales_by_model," "), x=Model, y =Total_sales_by_model), hjust=0.5, vjust=-0.2, fontface ="bold")

top_10_model_2021
```

Electricity

```{r}
electric_graph = ggplot(electric_price) + 
  geom_line(mapping = aes(x=Year, y =Annual_electricity), linetype = 1, color = "light blue", size =2, lineend = "round") +
  ylab("Annual Electricity Price (p/kWh)") +
  theme_bw() + theme(legend.title = element_blank())+
  geom_text(aes(label=ifelse(Year==2022, Annual_electricity," "), x=Year, y =Annual_electricity), hjust=1.5, fontface ="bold")

electric_graph
```

Petrol

```{r}
petrol_graph= ggplot(filtered_fuel, aes(x=Monthly_price, y=Pump_price))+
  geom_line(group=1, size =1.5, color = "light pink", alpha =0.5)+
  geom_point()+
  theme_minimal()+
  ylab("Price at pump (pence)")+ xlab ("Year")

petrol_graph
```

CPI

```{r}
CPI_graph = ggplot(CPI_Data_year, aes(x=Year, y=CPI_Index))+
  geom_point()+
  geom_smooth(color = "light blue", alpha =0.1)+
  theme_minimal()+
  ylab("CPI Index")

CPI_graph
```

## Analysis/ Insight

#### Annual car sales

-   There was a massive drop of annual car sales since 2020, likely due
    to COVID and post-COVID reduced production ability.
-   Expecting to have the number back to 2018-19 level, about 500,000
    difference.
-   Public are either driving old car or postpone buying their first car
-   **Potential for car sharing company to offer them experience of
    driving new car**

### Best selling car model

-   Difference between general model and specific model
-   Corsa is the top selling general model but not in the top 10
    specific model due to lots of difference sub-model under Corsa (e.g.
    petrol, diesel, electric)
-   6 out of 10 are hatchback, 3 are crossover SUV, 1 is medium SUV
-   2 is full electric, 5 are mild hybrid

### CPI index

-   reach record high in last 20 years and expected to be remained high
    in the next year
-   new car price will be high

### Electric and petrol price

-   Both reach record high level since 2010
-   Cost for running car will be high in the foreseeable future
-   No correlation shown between fuel price and car sales

## Conclusion

The overall number of car sales is reduced in the last 2 years due to
logistic, manufacturing problems. The car price will likely remained
high in the near future. There is a potential to offer people with new
car experience. As the hatchback and small/crossover SUV are the most
popular car model in the market and electric and mild-hybrid cars are
also gaining momentum. These car type should be considered when setting
the fleet initially. The running cost of the fleet will also be high as
the fuel price remained record high. Thanks
